Gametic Selection, Sex Ratio Bias, and Transitions Between Sex Determination Systems

### 2016-09-15 LaTeX/Turnover.tex
fixed male female inequality

%weak selection
Assuming weak selection, we can solve for the invasion fitness of the new sex -determining alleles explicitly, giving
\begin{equation}
\lambda_{m,XY} \approx 1 + V_A S_A C_m,
\end{equation}
where $V_A = p_A(1-p_A)$ is the variance at the \textbf{A} locus, $S_A = s^f (t^m)^2 / \left[ r R \left( s^f + s^m \right)^2 \right]$ (INTERPRETATION?), $s^f$ and $s^m$ are the respective selection coefficients for $A$ in diploid females and males, $t^m$ is the selection coefficient for $A$ in gametes/gametophytes from males, and we've assumed equal dominance coefficients in the two sexes.

@@

For the neo-$W$ we have $C_W = [(2r(1-R)-R) s^f + (1-2r)R s^m]/2$.
In this case, even when the novel sex-determining locus is in looser linkage with the selected locus than is the ancestral sex-determining locus, $r<R$, a neo-$W$ can invade.
For example, with $R=1/2$ the neo-$W$ invades if there is any linkage between the ancestral sex-determining and selected loci ($r<1/2$), there is selection among gametes/gametophytes in males ($t^m\neq0$), and there is selection for or against $A$ in both males and females ($s^m s^f >0$) that is stronger in females than in males ($|s^f| > |s^m|$).

@@

One might think that when the sex of a zygote is under the control of its mother, there would be strong selection to balance the sex ratio among zygotes. 
However, we find that, as with offspring control, assuming weak selection the invasion fitness of a sex-determiner that is maternally controlled can be written
\begin{equation}
\lambda_{k,XY} \approx 1 + V_A S_A C_k,
\end{equation}
where
%\begin{equation}
%V_A = \frac{- \left[ t_m + h (s_{f}+s_{m}) \right] \left[ t_m + (1-h) (s_{f}+s_{m}) \right]}{(1-2h)^2(s_{f}+s_{m})^2}
%\end{equation}
%is the variance in $A$ ($V_A>0$ when the polymorphic resident equilibrium is stable), 
%with $h$ the dominance coefficient (equal in both sexes), 
$C_k$ is a term that depends on $k$.
%
%%k=0
%\begin{equation}
%C_0 = big \; ugly \; term
%\end{equation}
%
%%k=1
%\begin{equation}
%C_1 = \frac{\left[ (2r(1-R)-R) s^f + (1-2r)R s^m \right]}{2}
%\end{equation}
%
%%k=1/2
%\begin{equation}
%C_{1/2} = even \; bigger, \; uglier \; term
%\end{equation}
%
%When $k=1$ the maternal- and offspring-controlled results are the same, $C_1 = C_W$, since the mutation remains in females (and hence mutant daughters always have mutant mothers).
%However, with $k\neq1$ the results with maternal control differ from those with offspring control.
Of particular interest is $k=1/2$ (i.e., when the mother perfectly balances the sex ratio of her offspring).
When both recombination rates are small we have $C_{1/2} \approx R(s^m - s^f)/8 = \lim_{r \rightarrow 0} C_W/4$.
This implies that, at least under tight linkage, maternal ESD can invade whenever a neo-$W$ can.

>>>>

%weak selection
Assuming weak selection, we can solve for the invasion fitness of the new sex -determining alleles explicitly, giving
\begin{equation}
\lambda_{m,XY} \approx 1 + V_A S_A C_m,
\end{equation}
where $V_A = p_A(1-p_A)$ is the variance at the \textbf{A} locus, $S_A = s^f (t^m)^2 / \left[ r R \left( s^f + s^m \right)^2 \right]$ (\textcolor{red}{INTERPRETATION? LOSE THE R?}), $s^f$ and $s^m$ are the respective selection coefficients for $A$ in diploid females and males, $t^m$ is the selection coefficient for $A$ in gametes/gametophytes from males, and we've assumed equal dominance coefficients in the two sexes.

@@

For the neo-$W$ we have $C_W = [(2r(1-R)-R) s^f + (1-2r)R s^m]/2$.
In this case, even when the novel sex-determining locus is in looser linkage with the selected locus than is the ancestral sex-determining locus, $r<R$, a neo-$W$ can invade.
For example, with $R=1/2$ the neo-$W$ invades if there is any linkage between the ancestral sex-determining and selected loci ($r<1/2$), there is selection among gametes/gametophytes in males ($t^m\neq0$), and there is selection for or against $A$ in both males and females ($s^m s^f >0$) that is stronger in males than in females ($|s^m| > |s^f|$).

@@

One might think that when the sex of a zygote is under the control of its mother, there would be strong selection to balance the sex ratio among zygotes. 
However, we find that, as with offspring control, assuming weak selection the invasion fitness of a sex-determiner that is maternally controlled can be written
\begin{equation}
\lambda_{k,XY} \approx 1 + V_A S_A C_k,
\end{equation}
where
%\begin{equation}
%V_A = \frac{- \left[ t_m + h (s_{f}+s_{m}) \right] \left[ t_m + (1-h) (s_{f}+s_{m}) \right]}{(1-2h)^2(s_{f}+s_{m})^2}
%\end{equation}
%is the variance in $A$ ($V_A>0$ when the polymorphic resident equilibrium is stable), 
%with $h$ the dominance coefficient (equal in both sexes), 
$C_k$ is a term that depends on $k$.
%
%%k=0
%\begin{equation}
%C_0 = big \; ugly \; term
%\end{equation}
%
%%k=1
%\begin{equation}
%C_1 = \frac{\left[ (2r(1-R)-R) s^f + (1-2r)R s^m \right]}{2}
%\end{equation}
%
%%k=1/2
%\begin{equation}
%C_{1/2} = even \; bigger, \; uglier \; term
%\end{equation}
%
%When $k=1$ the maternal- and offspring-controlled results are the same, $C_1 = C_W$, since the mutation remains in females (and hence mutant daughters always have mutant mothers).
%However, with $k\neq1$ the results with maternal control differ from those with offspring control.
Of particular interest is $k=1/2$ (i.e., when the mother perfectly balances the sex ratio of her offspring).
When both recombination rates are small we have $C_{1/2} \approx R(s^m - s^f)/8 = \lim_{r \rightarrow 0} C_W/4$.
This implies that, at least under tight linkage, the invasion of maternal ESD is independent of $R$ (because $S_A\propto R^{-1}$) and can invade whenever a neo-$W$ can.
